---
title: "farsfunctions"
author: "Alexander Shemetev"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{farsfunctions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(farsfunctions)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



#  ----------------------------------------------------------------------------
#  COURSERA: BUILDING R PACKAGES
##  Peer-graded Assignment
  File: fars_functions.R
  Description of the package
  GNU GENERAL PUBLIC LICENSE Version 3, 29 June 2007
  Start: 27/June/2020
  End:   28/July/2020
#  ---------------------------------------------------------------------------

 Read file with FARS data

 The functions work with the data taken from the .csv file (including .bz2 archived), stored on disk,
 from the US National Highway Traffic Safety Administration's (US NHTSA) date Fatality Analysis Reporting System (FARS), the US nationwide census, containing the
 US Public Annual Data, regarding fatal accidents occured in motorized
 vehicle (like automobiles or bikes) traffic crashes.
``` {}
For more information, see the source links (US NHTSA offial site and Wikipedia):
https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars
https://en.wikipedia.org/wiki/Fatality_Analysis_Reporting_System
```
# Function fars_read
Import: From readr read_csv
 Main dependencies: dplyr tbl_df

 Parameters: filename A character string with the name of the file to read, see
   notes.

  A data frame with data readed from the csv file, or an error if the
   file does not exists.

  Open file related with the accident data files 2013-2015
  Function fars_read can read .csv files and .csv archived in .bz2 format and other
 formats compatible for the basic readr package for its function read_csv
  Note that in the accidents data set only years 2013, 2014, and 2015 exist.
 Parameters: filename - name of the file (may be written with the path). The filename argument/input
 for the function is a string with the name of the file to be opened
  Function fars_read returns a tibble MxN where M - number of observations; N - number of columns
 (50 for the accidents data files)
 Usage: fars_read(filename)
 Examples:
 ```{r eval=FALSE}
print("Don't run me")
 #dontrun{
 library(dplyr)
 library(readr)
 setwd("C:/Users/Your_Path/")
 dataf = fars_read("accident_2013.csv.bz2")
 dataf = fars_read("C:/Users/Your_Path/accident_2013.csv.bz2")
 head(dataf)
 #dontrun}
 ```
# Function: make_filename
Note: To generate file name use: \code{\link{make_filename}}
 For help use:make_filename}
 Details: In case if the file does not exist an error message is produced and execution stops.
 Open code:
 ```{r eval=FALSE}
fars_read <- function(filename) {
        if(!file.exists(filename))
                stop("file '", filename, "' does not exist")
        data <- suppressMessages({
                readr::read_csv(filename, progress = FALSE)
        })
        dplyr::tbl_df(data)
}
```

 Make data file name

 Make .csv data file name related to the given \code{year}
 The function does not check if the file is available.

 Parameters: year A string or an integer with the input \code{year}

  This function returns a string with the data file name for a given
   year, and the file path within the package.

 Examples:
 ```{r eval=FALSE}
 #dontrun{
 make_filename(2012)
 
 s accident_2012.csv.bz2
# For help use: 
?fars_read
# Open code:
make_filename <- function(year) {
        year <- as.integer(year)
        system.file("extdata",
                    sprintf("accident_%d.csv.bz2", year),
                    package = "fars",
                    mustWork = TRUE)
}
#dontrun}
```
# Function fars_read_years
Used to read the FARS years
 Ancillary function used by fars_summarize_years
 Primary function to the
 Parameters: years A vector with a list of years
 Main dependencies: dplyr mutate_
 Main dependencies: dplyr select_
 Main dependencies: magrittr "%>%"
  A data.frame including entries in data by month,
 or NULL if the year is not valid
 For help use: ?fars_read
 For help use: ?make_filename
 For help use: ?fars_summarize_years
 Examples:
  ```{r eval=FALSE}
 #dontrun{
 setwd("C:/Users/Your_Path/")
 fars_read_years(2013)
 
# Open code:
fars_read_years <- function(years) {
        lapply(years, function(year) {
          file <- make_filename(year)
                tryCatch({
                        dat <- fars_read(file)
                        dplyr::mutate_(dat,  year = "YEAR") %>%
                          dplyr::select_("MONTH", "year")
                }, error = function(e) {
                        warning("invalid year: ", year)
                        return(NULL)
                })
        })
}
#dontrun}
```
# Function fars_summarize_years
 Summarize FARS data by years

 This simple function summarizes yearly accidents data, by month
 Parameters: years A vector with a list of years to summarize by.

  A data.frame with number of accidents by years summarized by month
 Main dependencies: dplyr bind_rows
 Main dependencies: dplyr group_by_
 Main dependencies: dplyr summarize_
 Main dependencies: tidyr spread_
 Main dependencies: magrittr "%>%"
 For help use: ?fars_read_years
 Examples:
   ```{r eval=FALSE}
 #dontrun{
 setwd("C:/Users/Your_Path/")
 plot(fars_summarize_years(2013))
 fars_summarize_years(c(2013, 2015))

# Open code: 
fars_summarize_years <- function(years) {
  dat_list <- fars_read_years(years)
  dplyr::bind_rows(dat_list) %>%
    dplyr::group_by_("year", "MONTH") %>%
    dplyr::summarize_(n = "n()") %>%
    tidyr::spread_("year", "n")
}
 #dontrun}
```
# Function fars_map_state
 Display accidents map by state and year

 Displays a plot with a state map including the accidents location by year
 If the \code{state.num} is invalid the function shows an error
 Parameters: state.num An Integer with the State Code (alphabetical order)
 \tabular{cc}{
   \strong{State Code} \tab \strong{State Name}    \cr
   01 \tab  Alabama              \cr
   02 \tab  Alaska               \cr
   04 \tab  Arizona              \cr
   05 \tab  Arkansas             \cr
   06 \tab  California           \cr
   08 \tab  Colorado             \cr
   09 \tab  Connecticut          \cr
   10 \tab  Delaware             \cr
   11 \tab  District of Columbia \cr
   12 \tab  Florida              \cr
   13 \tab  Georgia              \cr
   15 \tab  Hawaii               \cr
   16 \tab  Idaho                \cr
   17 \tab  Illinois             \cr
   18 \tab  Indiana              \cr
   19 \tab  Iowa                 \cr
   20 \tab  Kansas               \cr
   21 \tab  Kentucky             \cr
   22 \tab  Louisiana            \cr
   23 \tab  Maine                \cr
   24 \tab  Maryland             \cr
   25 \tab  Massachusetts        \cr
   26 \tab  Michigan             \cr
   27 \tab  Minnesota            \cr
   28 \tab  Mississippi          \cr
   29 \tab  Missouri             \cr
   30 \tab  Montana              \cr
   31 \tab  Nebraska             \cr
   32 \tab  Nevada               \cr
   33 \tab  New Hampshire        \cr
   34 \tab  New Jersey           \cr
   35 \tab  New Mexico           \cr
   36 \tab  New York             \cr
   37 \tab  North Carolina       \cr
   38 \tab  North Dakota         \cr
   39 \tab  Ohio                 \cr
   40 \tab  Oklahoma             \cr
   41 \tab  Oregon               \cr
   42 \tab  Pennsylvania         \cr
   43 \tab  Puerto Rico          \cr
   44 \tab  Rhode Island         \cr
   45 \tab  South Carolina       \cr
   46 \tab  South Dakota         \cr
   47 \tab  Tennessee            \cr
   48 \tab  Texas                \cr
   49 \tab  Utah                 \cr
   50 \tab  Vermont              \cr
   51 \tab  Virginia             \cr
   52 \tab  Virgin Islands       \cr
   53 \tab  Washington           \cr
   54 \tab  West Virginia        \cr
   55 \tab  Wisconsin            \cr
   56 \tab  Wyoming
 }
 Parameters: year A string, or an integer, with the input \code{year}

 Main dependencies: maps map
 Main dependencies: dplyr filter_
 Main dependencies: graphics points
 For help use: ?fars_read
 ?make_filename
 References 2014 FARS/NASS GES Coding and Validation Manual
 Examples:
  ```{r eval=FALSE}
 \dontrun{
 fars_map_state(37, 2014)
 
# Open code:
fars_map_state <- function(state.num, year) {
  filename <- make_filename(year)
  data <- fars_read(filename)
  state.num <- as.integer(state.num)
  if(!(state.num %in% unique(data$STATE))) {
    stop("invalid STATE number: ", state.num)
  }
  data.sub <- dplyr::filter_(data, .dots = paste0("STATE==", state.num))
  if(nrow(data.sub) == 0L) {
    message("no accidents to plot")
    return(invisible(NULL))
  }
  is.na(data.sub$LONGITUD) <- data.sub$LONGITUD > 900
  is.na(data.sub$LATITUDE) <- data.sub$LATITUDE > 90
  with(data.sub, {
    maps::map("state", ylim = range(LATITUDE, na.rm = TRUE),
              xlim = range(LONGITUD, na.rm = TRUE))
    graphics::points(LONGITUD, LATITUDE, pch = 46)
  })
}
# dontrun}
```
# Function catfunction 
This function allows you to express your love of cats.
Asks question: Do you love cats? Defaults to TRUE.
Examples: 
```{r eval=FALSE}
 cat_function()
```


## Life axamples for using the code 
 Arabidopsis FARS data on accidents in 2015

 Data from a FARS dataset in
 50 variables and several thousand observations

 Type: data

 Usage: data(mydatagr15)
 Format An object of class \code{"cross"}; see \code{\link[qtl]{read.cross}}.

 Keywords: datasets

 References: Coursera datasets from the course Build R packages (available in August 2020)
  Examples:
 str(mydatagr15)
```{r eval=FALSE}
path = "C:/Users/Your_Path/"
subfolder = "inst/extdata/"
myfile = "accident_2015.csv.bz2"
mydata <- fars_read(paste0(path, subfolder, myfile))
head(mydata[ , 1:8], 10)
```

```{r import data, echo=FALSE}
fars_read1 <- function(filename) {
        if(!file.exists(filename))
                stop("file '", filename, "' does not exist")
        data <- suppressMessages({
                readr::read_csv(filename, progress = FALSE)
        })
        dplyr::tbl_df(data)
}
path = "C:/Users/Alex/Documents/R/HOMEPROJECTS/farsfunctions/"
subfolder = "inst/extdata/"
myfile = "accident_2015.csv.bz2"
mydatagr16 <- fars_read1(paste0(path, subfolder, myfile))
mydatagr17 <- mydatagr16[ , 1:8]
knitr::kable(head(mydatagr17, 10))
```


## Additional Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 
```{r eval=FALSE}
fars_map_state(39, 2015)
fars_map_state(29, 2015)
```

```{r, fig.show='hold', echo=FALSE}
fars_map_state1 <- function(state.num, year) {
  filename <- mydatagr16
  data <- mydatagr16
  state.num <- as.integer(state.num)
  if(!(state.num %in% unique(data$STATE))) {
    stop("invalid STATE number: ", state.num)
  }
  data.sub <- dplyr::filter_(data, .dots = paste0("STATE==", state.num))
  if(nrow(data.sub) == 0L) {
    message("no accidents to plot")
    return(invisible(NULL))
  }
  is.na(data.sub$LONGITUD) <- data.sub$LONGITUD > 900
  is.na(data.sub$LATITUDE) <- data.sub$LATITUDE > 90
  with(data.sub, {
    maps::map("state", ylim = range(LATITUDE, na.rm = TRUE),
              xlim = range(LONGITUD, na.rm = TRUE))
    graphics::points(LONGITUD, LATITUDE, pch = 46)
  })
}
fars_map_state1(39, 2015)
fars_map_state1(29, 2015)



# plot(1:10)
# plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Additional Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "Link to the developer's GitHub: click here."
([via](https://github.com/Alexandershemetev/farsfunctions))
